Alberta Mortality 2000 to 2022

### 2024-03-14 paper/paper.qmd
model section finished

\begin{align*} 
y_i|\lambda_i, \theta &\sim \mbox{NegativeBinomial}(\mu_i, \theta) \\
\log(\mu_i) & = \beta_0 + \beta_1 \times \mbox{cause}_i\\
\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
\beta_1 & \sim \mbox{Normal}(0, 2.5) \\
\theta & \sim \mbox{SomePrior}(.)
\end{align*}

@@

>>>>

\begin{align*} 
y_i|\lambda_i, \theta &\sim \mbox{NegativeBinomial}(\mu_i, \theta) \\
\log(\mu_i) & = \beta_0 + \beta_1 \times \mbox{cause}_i\\
\beta_0 & \sim \mbox{Normal}(0, 2.5)\\
\beta_1 & \sim \mbox{Normal}(0, 2.5) 
\end{align*}

@@

## Model justification

The choice of the Poisson and Negative Binomial models stems from their suitability for count data. The Poisson model is a natural starting point for modeling count data, assuming that each event occurs independently and the mean rate of occurrence is constant.

However, the Poisson model's restrictive assumption that the mean equals the variance often does not hold in real-world data, prompting the use of the Negative Binomial model. The Negative Binomial model relaxes the equal mean-variance assumption by introducing a dispersion parameter, making it well-suited for modeling overdispersed count data that is common in mortality records.

```{r}
#| message: false
#| echo: false
#| warning: false
#| label: tbl-meanvariance
#| tbl-cap: Showing the mean and variance of total deaths in each year from 2001 to 2022, in Alberta

total_deaths_stats <- alberta_cod %>%
  summarise(
    Mean_Deaths = mean(total_deaths),
    Variance_Deaths = var(total_deaths)
  )

# Use kable to create a table
kable_table <- kable(total_deaths_stats, format = "latex", digits = c(0, 3, 3),
                     col.names = c("Estimate", "Variance"), 
                     caption = "Mean and Variance of Total Number of Deaths Over 22 Years")

kable_table
```
